---
title: "Reproducible Research: Peer Assessment 1"
author: "Philip M. Woite"
date: "24 January 2016"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The text file *"activity.csv"* was read using *read.csv*, and subsequently "NA" observations were removed using *complete.cases* to form the data frames **activity_complete** and **activity_day**. The data frame **activity_day** is a summary of **activity_complete**, taking the sum of the steps taken on each of the unique days in the data frame **activity_complete**.

```{r}
require('ggplot2')

require('data.table')

require('dplyr')

require('tidyr')

require('lubridate')

require('doBy')

activity <- read.csv('activity.csv', header=TRUE)
activity_complete <- data.frame(activity[complete.cases(activity),])

activity_day <- data.frame(summaryBy(steps ~ date ,data = activity_complete, FUN = list(sum)))
```

## What is mean total number of steps taken per day?

Then plotting a histogram

```{r, echo=TRUE}
qplot(activity_day$steps, geom = "histogram")
```

Evaluating the mean and median of the steps taken per day:

```{r, echo=TRUE, eval=TRUE}
mean_steps <- mean(activity_day$steps.sum, na.rm=TRUE)
print(mean_steps)
median_steps <- median(activity_day$steps.sum, na.rm=TRUE)
print(median_steps)
```




## What is the average daily activity pattern?
A new data fram **activity_intervals** was created from the **activity_complete** data frame - again using *summaryBy* but now with the function being "mean" to take the average. This data is then plotted.

```{r}
activity_intervals <- data.frame(summaryBy(steps ~ interval, data = activity_complete , FUN = list(mean)))

ggplot(activity_intervals, aes(interval, steps.mean))+ geom_line()
```

Finding the interval containing the maximum steps:

```{r}
max_interval <- which.max(activity_intervals[,2])
print(activity_intervals[max_interval,])

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
