---
title: "Reproducible Research: Peer Assessment 1"
author: "Philip M. Woite"
date: "24 January 2016"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
The text file *"activity.csv"* was read using *read.csv*, and subsequently "NA" observations were removed using *complete.cases* to form the data frames **activity_complete** and **activity_day**. The data frame **activity_day** is a summary of **activity_complete**, taking the sum of the steps taken on each of the unique days in the data frame **activity_complete**.

```{r}
require('ggplot2')

require('data.table')

require('dplyr')

require('tidyr')

require('lubridate')

require('doBy')

activity <- read.csv('activity.csv', header=TRUE)
activity_complete <- data.frame(activity[complete.cases(activity),])

activity_day <- data.frame(summaryBy(steps ~ date ,data = activity_complete, FUN = list(sum)))
```

## What is mean total number of steps taken per day?

Then plotting a histogram

```{r, echo=TRUE}
qplot(activity_day$steps, geom = "histogram")
```

Evaluating the mean and median of the steps taken per day:

```{r, echo=TRUE, eval=TRUE}
mean_steps <- mean(activity_day$steps.sum, na.rm=TRUE)
print(mean_steps)
median_steps <- median(activity_day$steps.sum, na.rm=TRUE)
print(median_steps)
```




## What is the average daily activity pattern?
A new data fram **activity_intervals** was created from the **activity_complete** data frame - again using *summaryBy* but now with the function being "mean" to take the average. This data is then plotted.

```{r}
activity_intervals <- data.frame(summaryBy(steps ~ interval, data = activity_complete , FUN = list(mean)))

ggplot(activity_intervals, aes(interval, steps.mean))+ geom_line()
```

Finding the interval containing the maximum steps:

```{r}
max_interval <- which.max(activity_intervals[,2])
print(activity_intervals[max_interval,])
```

## Imputing missing values
The number of missing values in *steps* in the **activity** data frame:

```{r}
missing_steps <- sum(is.na(activity$steps))
print(missing_steps)
```

To imput missing values, it was chosen to substitute *NA* values with the mean value in the interval of each of the *NA* values.

```{r}
activity_filled <- activity

activity_filled$steps <- ifelse(is.na(activity_filled$steps) == TRUE, activity_intervals$steps.mean, activity_filled$steps)

activity_filled_day <- data.frame(summaryBy(steps ~ date ,data = activity_filled, FUN = list(sum)))

```

Plotting and calculating the mean and median of the **activity_filled_day** data frame:

```{r, echo=TRUE}
qplot(activity_filled_day$steps, geom = "histogram")
```

Evaluating the mean and median of the steps taken per day:

```{r, echo=TRUE, eval=TRUE}
mean_filled_steps <- mean(activity_filled_day$steps.sum, na.rm=TRUE)
print(mean_filled_steps)
median_filled_steps <- median(activity_filled_day$steps.sum, na.rm=TRUE)
print(median_filled_steps)
```

The mean and median totals calculated now and in the earlier section differ. This difference may be ascribed to the substitution of average values skewing the median to the average. The second difference that was observed is that the mean and median values are now the same.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
activity_weekdays <- mutate(activity_filled_day, my_weekday = factor(weekdays(as.POSIXct(activity_filled_day$date))))

